---
title: "Practical Machine Learning"
author: "Sudarshan Dodda"
date: "December 26,2015"
output: html_document
---

## Background

Using devices such as JawboneUp, NikeFuelBand, and Fitbitit is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Preparing the data
Load packages, set caching

```{r}
library(RCurl)
library(caret)
library(corrplot)
library(ggplot2)
library(doMC)
knitr::opts_chunk$set(cache=TRUE)
```

##Getting Data

```{r}
# URL of the training and testing data
train.url ="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

x <- getURL(train.url, ssl.verifypeer = FALSE)
y<- getURL(test.url, ssl.verifypeer = FALSE)
train <- read.csv(textConnection(x))
test <- read.csv(textConnection(y))
dim(train)
dim(test)
```

##Replace NA values with 999
```{r}
test[is.na(test)]<-999
train[is.na(train)] <- 999
```

##Removing columns which has Id and Timestamp columns from both train and test
```{r}
train$X <- NULL
train$user_name <-NULL
train$raw_timestamp_part_1 <-NULL
train$raw_timestamp_part_2 <- NULL
train$cvtd_timestamp <- NULL

test$X <- NULL
test$user_name <-NULL
test$raw_timestamp_part_1 <-NULL
test$raw_timestamp_part_2 <- NULL
test$cvtd_timestamp <- NULL
```

##Checking the factor levels in classe
```{r}
str(train$classe)
levels(train$classe)
```


##Assigning classe to outcome
```{r}
outcome <- train[,"classe"]
```

##Remove 'classe' from train
```{r}
train$classe <- NULL
```

##Preprocessing
####Check for Zero Variance
```{r}
nzv.train = nearZeroVar(train, saveMetrics=TRUE)
nzv.train$nzv
```

##Removing the columns which has Near Zero Variance(nzv) is TRUE from both train and test data sets
```{r}
train_new <- train[c(rownames(nzv.train[nzv.train$nzv ==FALSE,]))]
test_new <- test[c(rownames(nzv.train[nzv.train$nzv ==FALSE,]))]
```

##Plot of relationship between features and outcome
Plot the relationship between features and outcome. From the plot below, each features has relatively the same distribution among the 5 outcome levels (A, B, C, D, E).


```{r}
featurePlot(train_new,outcome, "strip")
```


##Plot of correlation matrix
Plot a correlation matrix between features.
A good set of features is when they are highly uncorrelated (orthogonal) each others. The plot below shows average of correlation is not too high, so I choose to not perform further PCA preprocessing.

```{r}
corrplot.mixed(cor(train_new), lower="circle", upper="color", 
               tl.pos="lt", diag="n", order="hclust", hclust.method="complete")
```

## combining outcome with train_new dataset
```{r}
train_combined <- cbind(train_new,outcome)
```

## set random seed, for reproducibility 
```{r}
set.seed(1234)
```

##Data Splitting
```{r}
inTrain <- createDataPartition(y=train_combined$outcome,p=0.8, list=FALSE)
trainData <- train_combined[inTrain,]
validationData <- train_combined[-inTrain,]
```

```{r}
## set my cores 
registerDoMC(cores = 8)
```

## Model Building
```{r}
#fitcontrol <- trainControl(method = 'repeatedcv',number = 5,repeats = 3)
fitcontrol <- trainControl(method = "none")
gbmmodel <- train(outcome ~.,data = trainData,trControl = fitcontrol,
                  tuneGrid = expand.grid(n.trees = 150,
                                         interaction.depth = 3,shrinkage = 0.1),
                                         method = 'gbm',metric="Kappa")
```

##Top 20 features
```{r}
gbmImp <- varImp(gbmmodel,scale = FALSE)
plot(gbmImp,top =20)
```

##Predictions
```{r}
predictions <- predict(object = gbmmodel,validationData)
```

##Cross-validation
```{r}
confusionMatrix(validationData$outcome,predictions)
```

##Final Model Testing
```{r}
results <- predict(gbmmodel,newdata=test_new)
print(as.data.frame(results))
```
